### 2013-11-21 research/paper/Theory/mathematical_model.tex
Committing comments on abstract and introduction.
Committing mathematical_model.tex

\section{Background theory}

\subsection{Mathematical model (Quasi-steady )}

 One of the widely used mathematical model to predict the system response under galloping is the Quasi-steady state (QSS) model, incorporated by \cite{Parkinson1964} for a square cross section. The equation of motion of the body under galloping is given by Eq.\eqref{equationofmotion}. The forcing term $F_y$ is given by Eq.\eqref{lift equation}.

@@

\begin{equation}
\label{equationofmotion}
(m+m_a)\ddot{y}+c\dot{y}+ky=F_y
\end{equation}

\begin{equation}
\label{lift equation}
F_y=\frac{1}{2}\rho U^2\mathcal{A}C_y
\end{equation}

 In the QSS  model $C_y$ is determined by an interpolating polynomial based on the lift and drag data  for flow over a stationary body at various angles of attack. The order of the interpolation polynomial has varied from study to study. For  example a $7^{th}$ order polynomial was used in \cite{Parkinson1964} and $3^{rd}$ order polynomial was used in \cite{Barrero-Gil2009}. \cite{Ng2005} concluded that using a $7^{th}$ order polynomial is sufficient and a polynomial higher than that of $7^{th}$ order polynomial doesn't provides a significantly better result. Thus a $7 ^{th}$ order interpolating polynomial was incorporated in this present study.

\begin{equation}
\label{cy ploynomial}
C_y(\theta)=a_1\left(\frac{\dot{y}}{U}\right)+a_3\left(\frac{\dot{y}}{U}\right)^3+a_5\left(\frac{\dot{y}}{U}\right)^5+a_7\left(\frac{\dot{y}}{U}\right)^7
\end{equation}

@@

 It is expected that vortex shedding will be well correlated across the span and provide significant forcing at low \reynoldsnumber.\cite{Joly2012} introduced  an addition of a sinusoidal forcing function to the hydrodynamic forcing to model this. This enables to model to provide accurate predictions even at low mass ratios where galloping excitation is suppressed or not present. In this study, the forcing due to vortex shedding in low \reynoldsnumber cases is incorporated using a sinusoidal forcing function $F_0\sin{\omega_{s}t}$ added to the RHS of the equation. $\omega_{s}$ and $F_0$ represents shedding frequency and the maximum force due to shedding respectively. Thus, the final equation is given by Eq.\eqref{final_equation_motion}.

\begin{equation}
\label{final_equation_motion}
(m{+}m_a)\ddot{y}{+}c{+}\dot{y}{+}ky{=}\frac{1}{2}\rho U^2 \mathcal  {A} \Bigg(a_1\left(\frac{\dot{y}}{U}\right){+}a_3\left(\frac{\dot{y}}{U}\right)^3{+}a_5\left(\frac{\dot{y}}{U}\right)^5{+}a_7\left(\frac{\dot{y}}{U}\right)^7 \Bigg){+} F_0\sin{(\omega_s t)}
\end{equation}

This equation could be solved by time integration methods. In  this study  `Ode 45' routine in MATLAB was used to obtain the solutions. However in certain cases (low mass ratios) `Ode 15s' routine was used when Ode 45 had issues with stability due to the non-linear nature of the forcing.

@@

 The dissipated power due to the mechanical damping represents the ideal potential amount of harvested power output. Therefore the mean power output could be given by Eq.\eqref{power}.

 \begin{equation}
 \label{power}
P_{mean}=\frac{1}{T}\int_{0}^{T}(c\dot{y})\dot{y} dt
 \end{equation}

@@

The stationary data and the fluid-structure interaction (FSI) data were obtained using a high order spectral element routine to simulate the two-dimensional laminar flow. The Reynolds number was kept at 165 as it was pointed out by \cite{Sheard2009} and \cite{Tong2008} that the 3 dimensional transition for a square cylinder occurs at approximately Re=160. $F_0$ was kept at $0.4937$ which was obtained by scaling the value used by \cite{Joly2012} with the amplitude ratios of the lift forces obtained at the different Reynolds numbers.

Shedding frequency $\omega_s$, was set to $0.98$ which was obtained by performing a power spectral analysis of the stationary data at $0^0$. Stationary $C_y$ data were obtained at different angles of attack ranging from $0^0$ to $16^0$. The average power was obtained by using Eq.\eqref{power}

and the averaging was done over no less than 20 galloping periods. Predictions of power output at Re=22300 were obtained using coefficients for curve fitting $C_y$ (Table (\ref{table:cy-coefficients})) from \cite{Parkinson1964} in order to provide a comparison between high and low Reynolds numbers. $m^*$ was kept at 1163 for \reynoldsnumber=22300 (Similar as \cite{Parkinson1964}) and $m^*=20$ for \reynoldsnumber=165. These parameters  were used throughout this study unless otherwise specified.

 Simulations involving fluid structure interaction (FSI) were used to provide additional validation of the QSS model. The inlet was placed $20D$ while the outlet situated $60D$ away from the centroid of the body. The side boundaries were placed $20D$ away from the centroid of the body where $D$ was kept as unity throughout this study. The Naiver-Stokes equations were solved using an accelerated frame of reference using the previously mentioned code. A three-step time splitting scheme together with high-order Lagrangian polynomials were used to obtain the solution. The details of the method could be found in \cite{Thompson2006,Thompson1996a}. This code was incorporated in \cite{Leontini2011,Leontini2007a}  where it was employed in a fluid-structure interaction problems.

>>>>

\section{Background theory}
\label{sec:theory}

\subsection{The quasi-steady state (QSS) model}

The base quasi-steady state (QSS) model was first developed by \citet{Parkinson1964} for a square cross section. The equation of motion of the body  is given by 
\begin{equation}
\label{equationofmotion}
(m+m_a)\ddot{y}+c\dot{y}+ky=F_y,
\end{equation}
where the forcing term $F_y$ is given by
\begin{equation}
\label{lift equation}
F_y=\frac{1}{2}\rho U^2\mathcal{A}C_y.
\end{equation}

@@

In the QSS model, it is assumed that the force on the body at a given instantaneous angle of attack $\theta$ (defined in figure \ref{fig:setup_1}) is the same as the mean force on a static body at the same angle of attack. The instantaneous value of $C_y$ is therefore determined by an interpolating polynomial based on the lift and drag data  for flow over a stationary body at various $\theta$. Using the relationship between $\theta$ and the instantaneous transverse velocity of the body $\dot{y}$ shown in figure \ref{fig:setup_1}, $C_y$ can be written as a function of $\dot{y}$. The order of the interpolation polynomial used to define this functio has varied from study to study. For  example a $7^{th}$ order polynomial was used in \cite{Parkinson1964} and $3^{rd}$ order polynomial was used in \cite{Barrero-Gil2009}. \cite{Ng2005} concluded that using a $7^{th}$ order polynomial is sufficient and a polynomial higher than that of $7^{th}$ order doesn't provides a significantly better result. Thus a $7 ^{th}$ order interpolating polynomial is used in this present study. As a result, $C_y(\dot{y})$ is defined as
\begin{equation}
\label{cy ploynomial}
C_y(\theta)=a_1\left(\frac{\dot{y}}{U}\right)+a_3\left(\frac{\dot{y}}{U}\right)^3+a_5\left(\frac{\dot{y}}{U}\right)^5+a_7\left(\frac{\dot{y}}{U}\right)^7.
\end{equation}

@@

 It is expected that vortex shedding will be well correlated along the span and provide significant forcing at low \reynoldsnumber. \citet{Joly2012} introduced  an additional sinusoidal forcing function to the hydrodynamic forcing to model this. This enables the model to provide accurate predictions even at low mass ratios where galloping excitation is suppressed or not present. In this study, the forcing due to vortex shedding in low \reynoldsnumber\ cases is incorporated using a sinusoidal forcing function $F_0\sin{\omega_{s}t}$ added to the right-hand side of equation \ref{equationofmotion}. Here, $\omega_{s}$ and $F_0$ represent the angular vortex shedding frequency and the maximum force due to shedding respectively. Thus, the final equation for the modified QSS model is

\begin{equation}
\label{final_equation_motion}
(m{+}m_a)\ddot{y}{+}c{+}\dot{y}{+}ky{=}\frac{1}{2}\rho U^2 \mathcal  {A} \Bigg(a_1\left(\frac{\dot{y}}{U}\right){+}a_3\left(\frac{\dot{y}}{U}\right)^3{+}a_5\left(\frac{\dot{y}}{U}\right)^5{+}a_7\left(\frac{\dot{y}}{U}\right)^7 \Bigg){+} F_0\sin{(\omega_s t)}.
\end{equation}

This equation could be solved using standard time integration methods. In  this study the fourth-order Runge-Kutta scheme built in to the MATLAB routine `ode45' was generally used to obtain the solutions. Some low mass ratio cases used a solver modified for stiff problems, built into the `ode15s' routine in MATLAB.

@@

 The dissipated power due to the mechanical damping represents the ideal potential amount of harvested power output. Therefore, the mean power output can be given by
\begin{equation}
\label{power}
P_{mean}=\frac{1}{T}\int_{0}^{T}(c\dot{y})\dot{y} dt,
\end{equation}
where $T$ is the period of integration and $c$ is the mechanical damping constant.

It should be noted that this quantity is equal to the work done on the body by the fluid, defined as
\begin{equation}
\label{power_alt}
P_{mean}=\frac{1}{T}\int_{0}^{T}F_y\dot{y} dt,
\end{equation}
where $F_y$ is the transverse (lift) force.

These two definitions show two important interpretions of the power with respect to any energy production device. The first shows that power will be high for situations where the damping coefficient is high, and the transverse velocity is consistently high. The second shows that power will be high for situations where the transverse force and the body velocity are in phase.

@@

For the low \reynoldsnumber\ tests, \reynoldsnumber=165 was maintained as it was pointed out by \citet{Sheard2009} and \citet{Tong2008} that the three-dimensional transition for a square cylinder occurs at approximately \reynoldsnumber=160. $F_0$ was kept at $0.4937$ which was obtained by scaling the value used by \citet{Joly2012} with the amplitude ratios of the lift forces obtained at the different Reynolds numbers.

The angular vortex shedding frequency $\omega_s$, was set to $0.98$ which was obtained by performing a power spectral analysis of the stationary data at $0^\circ$. Stationary $C_y$ data were obtained at different angles of attack ranging from $0^\circ$ to $16^\circ$. The average power was obtained by using Eq.\eqref{power}, and the averaging was done over no less than 20 galloping periods. Predictions of power output at \reynoldsnumber=22300 were obtained using the coefficients for curve fitting $C_y$ (Table (\ref{table:cy-coefficients})) from \citet{Parkinson1964}, in order to provide a comparison between high and low Reynolds numbers. The mass ration $m^*$ was kept at 1163 for \reynoldsnumber=22300 (Similar as \citet{Parkinson1964}) and $m^*=20$ for \reynoldsnumber=165. These parameters were used throughout this study unless otherwise specified.

The stationary data and the fluid-structure interaction (FSI) data were obtained using a high-order spectral element routine to simulate the two-dimensional laminar flow.  Simulations involving fluid structure interaction (FSI) were used to provide additional validation of the QSS model. The inlet was placed $20D$ while the outlet situated $60D$ away from the centroid of the body. The side boundaries were placed $20D$ away from the centroid of the body where $D$ was kept as unity throughout this study. The Navier--Stokes equations were solved in an accelerated frame of reference attached to the moving body along with the body equation of motion given in equation \ref{equationofmotion}. A three-step time splitting scheme together with high-order Lagrangian polynomials were used to obtain the solution. The details of the method can be found in \cite{Thompson2006,Thompson1996a}. This code has been very well validated in fluid-structure interaction problems \citep{Leontini2011,Leontini2007a}.

### 2013-11-21 research/paper/nomenclature.tex
Committing mathematical_model.tex

\begin{tabular}{ll}
$a_1,a_3,a_5,a_7$ & coefficients of the polynomial to determine $C_y$ \\ 
$C_y$ & instantaneous force coefficient normal to the induced velocity \\
$F_y$ & force due to $C_y$  \\ 
$\rho$ & fluid density  \\
$m$ & mass of the body \\
$m_a$ & added mass \\
$c$ & damping constant/damping factor \\
$k$ & spring constant \\
$U_i$ & induced velocity \\
$\theta$ & induced angle of incidence \\
$U$ & freestream velocity \\
$y,\dot{y},\ddot{y}$ & transverse displacement, velocity and acceleration   \\
$A$ & displacement amplitude\\
$\mathcal{A}$ & cross sectional area\\ 
$F_0$& force due to shedding \\
$\omega_s$& vortex shedding frequency \\
$t$ & time \\
$P_{mean}$& mean power \\
$f=\frac{1}{2\pi}\sqrt{\frac{k}{m}}$& natural frequency of the system \\
$\omega_n= 2 \pi f$& natural frequency of the system  \\
$D$ & characteristic length of the body  \\
$m^*=\frac{m}{\rho \times \text{Volume of the body}}$& mass ratio \\
$U^*=\frac{U}{f \times D}$& reduced velocity  \\
$\zeta= \frac{c}{2 m \omega_n}$& damping ratio \\
$P_t$   & power transferred to the body by the fluid \\
$P_d$& power dissipated due to mechanical damping  \\
$Re$& Reynolds number  \\
$\theta= \tan^{-1}{(\frac{\dot{y}}{U})}$& instantaneous angle \\
& \\
& \\
\end{tabular}

>>>>

\begin{tabular}{ll}
$a_1,a_3,a_5,a_7$ & coefficients of the polynomial to determine $C_y$ \\ 
$C_y$ & instantaneous force coefficient normal to the induced velocity \\
$F_y$ & force due to $C_y$  \\ 
$\rho$ & fluid density  \\
$m$ & mass of the body \\
$m_a$ & added mass \\
$c$ & damping constant/damping factor \\
$k$ & spring constant \\
$U_i$ & induced velocity \\
$\theta$ & induced angle of incidence \\
$U$ & freestream velocity \\
$y,\dot{y},\ddot{y}$ & transverse displacement, velocity and acceleration   \\
$A$ & displacement amplitude\\
$\mathcal{A}$ & cross sectional area\\ 
$F_0$& amplitude of the oscillatory force due to shedding \\
$\omega_s$& vortex shedding frequency \\
$t$ & time \\
$P_{mean}$& mean power \\
$f=\frac{1}{2\pi}\sqrt{\frac{k}{m}}$& natural frequency of the system \\
$\omega_n= 2 \pi f$& natural frequency of the system  \\
$D$ & characteristic length of the body  \\
$m^*=\frac{m}{\rho \times \text{Volume of the body}}$& mass ratio \\
$U^*=\frac{U}{f \times D}$& reduced velocity  \\
$\zeta= \frac{c}{2 m \omega_n}$& damping ratio \\
$P_t$   & power transferred to the body by the fluid \\
$P_d$& power dissipated due to mechanical damping  \\
$Re$& Reynolds number  \\
$\theta= \tan^{-1}{(\frac{\dot{y}}{U})}$& instantaneous angle of attack\\
& \\
& \\
\end{tabular}
